---
title: "Astronaut Dataset Regression"
author: "Jeff Shamp"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: vignette
---


# Astronaut Missions 1960 - 2020

```{r, message=FALSE}
library(tidyverse)
library(tidymodels)
library(baguette)
```

Load data from the good people at RStudio

```{r, message=FALSE}
astronauts <- 
  readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-14/astronauts.csv')
```


```{r}
astronauts<- 
  astronauts %>%
  filter(hours_mission > 0 & in_orbit != "explosion") %>%
  select(name, sex, nationality, year_of_birth,
         year_of_selection, year_of_mission,
         military_civilian, occupation, in_orbit, 
         hours_mission) %>%
  mutate(occupation = str_to_lower(occupation), 
         occupation = case_when(
           str_detect(occupation, "^space  tourist$") ~ "space tourist", 
           str_detect(occupation, "space tourist") ~ "space tourist", 
           TRUE ~ occupation
            ), 
         occupation = str_replace_all(occupation, "other.*", "journalist")) %>%
  mutate(
        in_orbit = str_to_lower(in_orbit),
        in_orbit = str_remove_all(in_orbit, "[:digit:]"), 
        in_orbit = str_remove_all(in_orbit, "[:punct:]\\B"),
        in_orbit = case_when(
          str_detect(in_orbit, "^sts") ~ "sts", 
          str_detect(in_orbit, "^apollo") ~ "apollo", 
          str_detect(in_orbit, "^gemini") ~ "gemini", 
          str_detect(in_orbit, "^mir") ~ "mir", 
          str_detect(in_orbit, "^salyut") ~ "salyut",
          str_detect(in_orbit, "^saluyt") ~ "salyut",
          str_detect(in_orbit, "^skylab") ~ "skylab",
          str_detect(in_orbit, "^soyuz") ~ "soyuz",
          str_detect(in_orbit, "^ma$") ~ "mercury-atlas",
          TRUE ~ in_orbit)
        ) %>%
  mutate(hours_mission = log(hours_mission)) %>%
  na.omit()
```


```{r}
astronauts %>%
  mutate(year_of_mission = case_when(
      year_of_mission < (10*(year_of_mission %/% 10))+5 ~ (10*(year_of_mission %/% 10)),
      year_of_mission >= (10*(year_of_mission %/% 10))+5 ~  (10*(year_of_mission %/% 10))+5,
      TRUE ~ year_of_mission
  ), 
  year_of_mission = factor(year_of_mission)) %>%
  ggplot(aes(year_of_mission, hours_mission,
             fill=year_of_mission)) +
  geom_boxplot(alpha=0.5, show.legend = FALSE) +
  scale_fill_brewer(palette = "Paired") +
  theme_classic()
```


## Modeling

Building a regression model based on the selected data. The prediction will be the log of the hours of a given mission. 

```{r}
set.seed(9450)
astro_split<- initial_split(astronauts, strata = hours_mission, prop = .80)
astro_train<- training(astro_split)
astro_test<- testing(astro_split)
```

Adding a recipe for modeling. Updating the role of "id" and "name" to be ignored from the analysis and collapsing the lesser used occupations and orbiting modules as "Other". The `update_role` and `step_other` functions are nice additions. Also the `step_dummy` feature is a huge improvemtn of the fastdummies package. 

Creating the recipe will not return the results of the `step_other`, to see and edit the results of that function you have to `prep()` and `juice()` the `astro_recipe`. 

```{r}
astro_recipe<- 
  recipe(hours_mission ~ ., data = astro_train) %>%
  update_role(name, nationality, new_role = "ID") %>%
  step_other(occupation, in_orbit, other="Other") %>%
  step_dummy(all_nominal(), -has_role("ID")) 

astro_recipe
```


```{r}
astro_wf<- 
  workflow() %>%
  add_recipe(astro_recipe) 
  

astro_wf
```


```{r}
tree_model<- 
  bag_tree() %>%
  set_engine("rpart", times = 30) %>%
  set_mode("regression")

tree_model
```

```{r}
mars_model<- 
  bag_mars() %>%
  set_engine("earth", times=30) %>%
  set_mode("regression")

mars_model
```


```{r}
tree_results<- 
  astro_wf %>%
  add_model(tree_model) %>%
  fit(astro_train)

mars_results<- 
  astro_wf %>%
  add_model(mars_model) %>%
  fit(astro_train)
```


```{r}
tree_results
mars_results
```

Hours of the mission is the largest predictor by an order of magnitude for both models. The tree model uses more the other predictors, whereas the mars model seems to collapse more the predictors importance. 

We can now make predictions and tie those predictions to the testing dataframe. 

```{r}
test_results<- 
  astro_test %>%
  bind_cols(predict(tree_results, astro_test)) %>%
  rename(pred_tree = .pred) %>%
  bind_cols(predict(mars_results, astro_test)) %>%
  rename(pred_mars = .pred)
```


```{r}
test_results %>%
  metrics(hours_mission, pred_tree)
```


```{r}
test_results %>%
  metrics(hours_mission, pred_mars)
```

```{r}
new_astronauts <- crossing(
  in_orbit = fct_inorder(c("iss", "sts", "mir", "Other")),
  military_civilian = "civilian",
  occupation = "Other",
  year_of_mission = seq(1960, 2020, by = 5),
  name = "ID", 
  year_of_selection = seq(1960, 2015, by = 5),
  year_of_birth = seq(1920, 1990, by =5), 
  sex = "male", 
  nationality = "U.S.S.R/Russia"
) %>%
  filter(
    !(in_orbit == "iss" & year_of_mission < 2000),
    !(in_orbit == "mir" & year_of_mission < 1990),
    !(in_orbit == "sts" & year_of_mission > 2010),
    !(in_orbit == "sts" & year_of_mission < 1980)
  )

```


```{r}
new_astronauts %>%
  bind_cols(predict(mars_results,  new_astronauts)) %>%
  ggplot(aes(year_of_mission, .pred, color = in_orbit)) +
  #geom_point(size = 2, alpha =0.5) +
  geom_jitter(alpha=0.4)
  labs(
    x = NULL, y = "Duration of mission in hours (predicted, on log scale)",
    color = NULL, title = "How did the duration of astronauts' missions change over time?",
    subtitle = "Predicted using bagged decision tree model"
  )
```






